An efficient composition of bidirectional programs by tupling and lazy updates

### 2019-11-30 paper/6.related-work.tex
revised related work.

\section{Related Work}

\subsubsection*{BX Languages and Implementations}

There are many BX languages \cite{Buchmann:2018:BFI:3362232.3362263,Ko:2016:BFV:2847538.2847544,LeblebiciAS14,Samimi-Dehkordi18,Cicchetti2011,Hinkel:2019:CPB:3318595.3318634}
%(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)
based on several application scenarios and BX lenses works \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487}. The semantics and correctness of bidirectional programs has been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood. Anjorin et. al. introduces the first benchmark for BX languages and compared them \cite{Anjorin2019}. However, the improvement for practical implementation of BX languages is still missing. As a first step, we study different evaluation methods to optimized their evaluation. We can say this paper will be first attempt to improve efficiency of BX composition evaluation. In this paper we focus on a BX language BiGUL \cite{Ko:2016:BFV:2847538.2847544, Ko:2017:ABB:3177123.3158129}, we compare BiGUL's method (in Section 2) and our methods (in Section 3 and 4). From experimental results of left associative BX composition programs, we know that our memoization methods are faster than the original BiGUL's evaluation method.
In this paper we focus on BiGUL, however our methods can be applicable to other BX languages which have similar problems.

@@

\subsubsection*{Optimization Techniques}

Memoization \cite{Bellman:2003:DP:862270,MICHIE1968} is well-known technique to reduce extra evaluation. In our case, simple application of memoization $put_m$ works well for many BX programs.
Another way to reduce extra evaluation is fusion \cite{Wadler:1988:DTP:80099.80104}. However, fusion of BX programs are not considered well \textcolor{red}{(really?)}. Therefore we did not use this way.
  Instead of that, we introduce another way that uses tupling \cite{Fokkinga90}, lazy update and lazy computation \cite{Henderson:1976:LE:800168.811543, Hudak:2007:HHL:1238844.1238856}. Tupling is often used for reducing common evaluations, but we use it for reducing the size of complements.
  Lazy evaluation is also a well-known and widly-used optimization technique to reduce the redundant evaluation. Because we use OCaml for implementation, we need to introduce this explicitly. If we use Haskell, a lazy evaluated language, we do not need to introduce them.

>>>>

\section{Related Work} \label{sect:related}

%\subsubsection*{BX Languages and Implementations}

Since the pioneer work of Lens \cite{Foster:2007:CBT:1232420.1232424}, many BX languages have been proposed \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487,Buchmann:2018:BFI:3362232.3362263,Ko:2016:BFV:2847538.2847544,LeblebiciAS14,Samimi-Dehkordi18,Cicchetti2011,Hinkel:2019:CPB:3318595.3318634}.
%(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)
%based on several application scenarios and BX lenses works \cite{}.
Although much progress has been made on the semantics and correctness of BX programs for the past years, as far as we are aware, little work has been done on optimization of BX programs \cite{Horn:2018:IRL:3243631.3236769}. Anjorin et. al. introduces the first benchmark for BX languages and compared them \cite{Anjorin2019}, but a systematic improvement for practical implementation of BX languages is still missing. This paper shows the first attempt of improving efficiency of BX composition evaluation.

The baseline of this work is the BX language BiGUL \cite{Ko:2016:BFV:2847538.2847544, Ko:2017:ABB:3177123.3158129}, and we compare BiGUL's method (in Section~\ref{sect:minbigul}) with our methods (in Sections~\ref{sect:minbigulm} and \ref{sect:xpg}). From experimental results of left associative BX composition programs, we can see that our memoization methods are faster than the original BiGUL's evaluation method.
While we focus on BiGUL, our methods are general and should be applicable to other BX languages.

@@

%\subsubsection*{Optimization Techniques}

Our work is related to many known optimization methods for unidirectional programs. Memoization \cite{Bellman:2003:DP:862270,MICHIE1968} is a technique to avoid repeated redundant computation. In our case, we show that two specific memoization methods can be used for bidirectional programs.
To deal with inefficiency due to compositions, many fusion methods have been studied \cite{Wadler:1988:DTP:80099.80104} to merge a composition of two (recursive) programs into one. However, under the context of bidirectional programs, we need to consider not only compositions of recursive programs but also compositions inside a recursive program (as we have seen in \mbox{\it bfoldr}). This paper focuses on the composition inside a recursion, where compositions are produced dynamically at runtime. We tackled the problem by using tupling \cite{Fokkinga90}, lazy update and lazy computation \cite{Henderson:1976:LE:800168.811543, Hudak:2007:HHL:1238844.1238856}.
